The transboundary nature of the world’s exploited marine species

### 2020-10-14 Scripts/Manuscript_Draft.Rmd
included proofs

Regulatory boundaries and species distributions often do not align. This is especially the case of marine species crossing multiple Exclusive Economic Zones (EEZs). Such movements represent a challenge for fisheries management as policies tend to focus at the national level, yet collaborations are needed to maximize long-term ecological, social and economic benefits of shared marine species. Here, we combined species distributions and the spatial delineation of EEZs to identify the number of commercially exploited marine species that are shared between neighboring nations. We found that 67% of the species analyzed are transboundary (n = 633). Between 2005 and 2014, fisheries targeting these species within global-EEZs caught on average 48 million tonnes per year, equivalent to an average of USD 77 billion in annual fishing revenue. For select countries, over 90% of their catch and economic benefits were attributable to a few shared resources. Our analysis suggests that catches from transboundary species are declining more than those from non-transboundary species. Our study has direct implications for managing fisheries on transboundary species, highlighting the need for strengthened effective international cooperation.

@@

Distributions of marine species around the world are not constrained by human-made boundaries; rather they are shaped by biotic and abiotic factors as well as species' evolutionary history [@Hutchinson:1957nt; @Nelson:2016kk]. A species can be widely distributed (cosmopolitan) or endemic [@Nelson:2016kk]. Fisheries management is predicated on the definition of “stocks”, delineated, in most cases by human-made spatial boundaries that often do not correspond to biologically-meaningful population units [@Song:2017iua; @FredstonHermann:2018kp]. The **development** of Exclusive Economic Zones (EEZs) under Part V of the United Nations Convention on the Law of the Seas (UNCLOS) in the early 80s, for instance, extended political boundaries from 12 to 200 nautical miles to give coastal nations property rights over marine resources [@osthagen:2020hd]. However, EEZ boundaries cut across the distribution of many species, creating shared stocks between nations [@UnitedNations:1986tl]. Shared stocks can be classified into three non-exclusive categories; "transboundary" - stocks that cross the EEZs of two or more coastal states; "straddling" - stocks that cross neighboring EEZs and the adjacent high seas; and "highly migratory" - stocks that cross non-neighboring EEZs and the high seas (mainly tunas) [@Munro:2004th]. Our study focuses on the "transboundary" nature of shared species exploited by fisheries operating within EEZs.

Theories and empirical evidence have shown that fisheries targeting resources that straddle political boundaries complicate fisheries management and potentially reduce the effectiveness of policies to achieve their stated objectives [@Munro:2002uf; @Englander:2019dq]. For instance, climate-driven changes in species distributions have led to conflicts between nations driven, at least partly, by changes in the proportion of captures [@Munro:2002uf], quota allocation, and fishery newcomers [@Spijkers:2017ij]. Moreover, climate change is likely to exacerbate such conflicts and presents new challenges for political relations between neighboring countries [@Pinsky:2018cb] and fisheries management [@Miller:2013iv]. Therefore, having an accurate understanding of the distribution and scale of transboundary and straddling fish stocks as well as associated fisheries is important to inform their sustainable management, particularly under climate change.

Forty years after the formal adoption of UNCLOS and the subsequent 1995 United Nations Fish Stock Agreement for the cooperation on the management and conservation of straddling **and highly migratory** resources [@UnitedNations:1995fs], accurate estimates of the number of exploited marine species shared by neighboring nations are still unavailable. An informed guess based on limited biogeographical data, suggested that there were approximately 500 to 1500 exploited transboundary stocks [@Caddy:1997ue]. A recent literature review included 344 shared taxa and their historical contributions to fisheries [@Teh:2015gd]. However, these studies did not consider species' biogeography in the quantification of transboundary stocks. Here, we aim to estimate the number of exploited marine species shared by neighboring countries and determine their contribution to global and regional catches aas well as fisheries revenue. Moreover, we categorize species according to their catch trends while identifying the differences among species based on habitat preference. We hypothesize that the methodological constraints of previous studies resulted in an underestimation of the number of transboundary stocks and their contribution to global catch and revenue.

We overlaid the known distribution of 938 commercially valuable marine species responsible for an average of 96.5% of global EEZ catches between 2005 and 2014, and 280 EEZs of 198 coastal countries (See methods). We define a ‘stock’ unit as a species in an EEZ, instead of a genetically or morphological distinct units [@Nelson:2016kk], due to a lack of such biological information for almost all the species included in this study [@Teh:2015gd]. While we acknowledge that a species could have multiple stocks within an EEZ, many fisheries within a country or EEZ are managed at the species instead of stock level (or even as groups containing multiple species). For example, shrimp (e.g. *Litopenaeus stylirostris* or *Farfantepenaeus californiensis*) in the Pacific coast of Mexico [@DiarioOficialdelaFederacionDOF:2018wt] and hammerhead shark (*Sphyrna zygaena*) in Peru [@MAP:2017uf] are managed at the species level in both cases, yet include multiple populations exploited by each country’s fisheries. **Moreover, recent research shows connectivity between fish stocks through larval dispersal [@Ramesh:2019va] and adult migration [@Levin:2018in; @Popova:2019fs; @Dunn:2019ca], although considerable level of uncertainty exists at different life stages [@Kaplan:2016hs; @Archambault:2016ie].** For this analysis we only considered shared species between neighboring EEZs, rather than the species’ extended distribution (e.g. we did not include the high seas). We rely on multiple data sources including occurrence, distribution models and catch data and only consider a species to be present in a grid cell if all data sources showed positive occurrence (see Methods).

@@

We identified 633 exploited transboundary species worldwide (67.5% of the 938 species analyzed), almost double previous estimates [@Teh:2015gd]. Between 2005 and 2014, national fleets targeting these transboundary species within EEZs caught an annual average of 48.5 million tonnes, representing 82.3% of EEZ-derived catches reconstructed by the *Sea Around us* at the species level (Fig. 1a). These catches generated a yearly average of USD 77,591 million in fishing revenue (78.5% of the global fishing revenue) over the same time period. Our findings are considerably higher than the previous estimates of 34.2 million tonnes in catches and USD 40,187 million (in 2019 value) in fishing revenue from shared stocks [@Teh:2015gd].When we re-estimated the transboundary species’ catches using data consistent with those used in previous studies (FAO global reported data in year 2006), the 633 transboundary species identified here accounted for 40.4 million tonnes of annual catches (i.e., 18% higher than previously estimated). Our results suggests that the contribution of transboundary species to global catch and fishing revenue might have been previously underestimated due to an incomplete understanding of the transboundary nature of marine fished species. The 305 non-transboundary species (termed as ‘discrete’ species here, see methods - Determining transboundary species trait) accounted for a much smaller proportion of total catch and revenue; 2.8 million tonnes and USD 4,282 million annually respectively, on average, between 2005 and 2014. These results underscore the importance of transboundary species at the global level.

@@

In many cases, according to our categorization criteria, a transboundary species **can be distributed** in multiple EEZs but only counted as transboundary in a subset of EEZs (see Methods). For example, the distribution of Peruvian anchoveta (*Engraulis ringens*) spans the EEZs of Peru, Chile, and Ecuador. However, our study only considered the stocks in Peru and Chile as transboundary [@Cashion:2018cg]. Anchoveta in Ecuadorian waters only include a small proportion of the shared distribution range (spatial threshold between Ecuador and Peru < 10%; see methods - Criteria 3) and thus do not meet our criteria for consideration as transboundary stock. **A situation similar** to the example of Peruvian anchoveta in Ecuadorian waters is common amongst the transboundary species identified in this study. Overall, 590 of the 633 transboundary species have stocks in EEZs that do not meet our criteria for consideration as transboundary stock. The annual average contribution from all excluded stocks of transboundary species to fisheries (e.g., Ecuador’s anchoveta catch) between 2005 and 2014 was 10.8 million tonnes, representing USD 19,243 million in fishing revenue over the same time period.

@@

Peru and Russia, having the two largest fisheries by total catch in the world [@FAO:2018KK], each responsible for over 5.8 million tonnes annually, on average, between 2005 and 2014  (Supplementary Fig. S1). Peru's catches consist mainly of Peruvian anchoveta (*Engraulis ringens*) and account for 79% of the national transboundary species production. Peru and Chile recently signed an agreement to work towards standardized stock assessments through coordinated management of the southern anchoveta stock [@UNDP:2016wp]. A similar management agreement was signed by Russia, Japan and the USA over shared Alaskan pollock (*Theragra chalcogramma*) in the Bering Sea in 1988 [@NOAA:gX6CtUTk]. Transboundary species also make large contributions to fisheries in Eastern Asia (one-way ANOVA; *Df*(16,158) = 2.265, *p* = 0.005, $\alpha$ = 0.05). China, the world's top fish producer [@FAO:2018KK], obtains one third (5.1 million tonnes) of its total catches from transboundary species, followed by Japan (1.8 million tonnes) and South Korea (1.06 million tonnes). Differences in the regional importance of transboundary fisheries can also be illustrated in terms of catch-revenue over the area ($km^2$) of the EEZ (Fig. 2). As an example, the aggregated EEZ area for all Northern European countries where transboundary species are present was 3.3 million $km^2$, the $6^{th}$ smallest of the 17 groups analyzed (Supplementary Table S1). However, it had the second highest average revenue (USD 26.1 thousand per $km^2$) and the highest average catch (19.9 tonnes per $km^2$) of transboundary species per EEZ area between 2005 and 2014. At the country level, countries from Western Europe accrued significantly more revenue from transboundary fisheries per $km^2$ than any other country (one-way ANOVA, *Df*(16,165) = 3.267, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* $\leq$ 0.05; Fig. 2).

@@

![Figure 2. Weighted benefits of transboundary species by $km^{2}$ and UN sub-region. a) Revenue in thousand 2019 USD. b) Catch in tonnes. Points are sub-region averages $\pm$ s.d. by country. Figure created using R version 3.5.2. Abbreviations as in Figure 1.](../Figures/Fig2.png)

We determined the catch trend of each species within each EEZ, classifying them as increasing (Category A), constant (Category B) or decreasing (Category C) (Fig. 3). While previous studies have demonstrated that catch trends can be used to infer whether a stock is healthy, re-building, over-exploited or collapsed [@Kleisner:2011wn], several factors can influence stock status. Our intention here is to examine where the catch trends of transboundary species differ from non-transboundary species between 2005 and 2010 relative to historic catch since 1951 (See methods). We found significant differences for all catch trend categories for transboundary species (one-way ANOVA, *Df*(2,459) = 47.94, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* $\leq$ 0.001), and no significant differences in catch trend categories for discrete species (one-way ANOVA, *Df*(2,106) = 1.885, *p* = 0.157; $\alpha$ = 0.05). We also found significant differences in catch trends when directly comparing transboundary to discrete species categories (MANOVA, *Df*(2,459) = 19.001, *p* < 0.001). Overall, transboundary species only targeted by one country are generally less likely to have a decreasing catch trend compared to instances where the shared species is fished by neighboring countries (Supplementary Table S2).

@@

![Figure 3. Number of transboundary species by catch trend and EEZ. Category A, Increasing; Category B, Constant; Category C, Decreasing. “No Category” reflects species with less than 10 years of catch data and/ or less than 5 consecutive years of catch data. Figure created using R version 3.5.2.](../Figures/Fig3.png)

Empirical analysis suggests that in most cases, management of transboundary species will yield better outcomes in terms of fish catches when nations cooperate [@Munro:2002uf]. Yet, cooperation can be a complex process [@Jensen:2015cf] with specific cases where joint management might not be the best strategy [@Munro:2015df]. Examples of successful joint management include agreements between Norway and Russia over Atlantic cod (*Gadus morhua*) [@Eide:2013dha] and Namibia and South Africa over hake (*Merluccius spp*) [@Sumaila:2003vw]. Lack of collaboration over shared stocks may threaten stock sustainability, reduce the future profitability potential of the fishery, and result in conflict between neighboring nations [@Clark:1980js; @Spijkers:2017ij].

<!-- Europe has the higest rate of transboundary species per $km^2$ (Figure 2) which is consistent with their Common Fisheries Policy to manage the EEZ as a whole. However, despite some cases of successfull fisheries management, excessive quotas, poorly implemented fishing subsidies, and policy-related issues have failed to control fishing effort threating the sustainabillity of most european stocks [@Khalilian:2010bk].  -->

Transboundary fisheries are important to a number of countries with notorious fisheries-related conflicts, including Canada, the USA, Russia and the European Union (EU) [@Spijkers:2019cz]. For example, since 2007, the EU, Norway, Iceland, and the Faroe Islands (Denmark) have been at odds over the size and relative allocation of total allowable catches for Atlantic mackerel (*Scomber scombrus*) due to the species climate-driven shift in distribution [@Spijkers:2017ij]. Atlantic Mackerel is a transboundary species that straddles into the high seas. Amongst the countries involved in the 2007 fisheries dispute, Atlantic mackerel contributed an annual average (between 2005 and 2014) 598.4 thousand tonnes (8.19%) in total catchs and USD 850.2 million (7.17%) in total fishing revenue. Climate change is expected to continue changing the distribution and shared proportion of fish stocks among countries, resulting in the emergence of new transboundary species [@Pinsky:2018cb], and disappearance of some species from EEZs [@Oremus:2020id]. Exploring the detailed effects of climate change on the distribution of shared species is key to the development of local adaptation methods that can anticipate negative impacts to sustainability. For example, understanding how climate change will modify the proportion of transboundary species shared by neighboring EEZs, the time-frame over which such changes will happen [@PalaciosAbrantes:2020vt], and the economic consequence of such effects can inform the development of more climate-resilient international treaties improving fisheries management [@Sumaila:2020tp].

Most marine fish species occur in tropical and subtropical waters around the world [@Nelson:2016kk; @Reygondeau:2019uh], from highly migratory species associated with pelagic-oceanic ecosystems like tunas (*Thunnus sp.*), to less mobile reef-associated species like greater amberjack (*Seriola dumerili*), and species found in demersal ecosystems like gilthead seabream (*Sparus aurata*). Species associated to pelagic-oceanic ecosystems are the only group whose EEZ range (i.e., the number of EEZs where the species occur as transboundary) was significantly different from other groups (one-way ANOVA; *DF*(5,597) = 53.82, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05) with a median of 40 EEZs per species. The median for species of all other ecosystem preferences is close to, or less than, 20, as many of these species have a narrower distribution or are less mobile (Fig. 4a). This result is likely an effect of the broad distribution that many of the species with preference for pelagic-oceanic ecosystems have. Many pelagic-oceanic species are highly migratory and straddle across EEZs while crossing the high seas. Thus, the number of non-neighboring EEZs sharing a highly migratory species can be over 100, as in the case of bigeye tuna (*Thunnus obesus*) (Fig. 4a). Due to their vast migration patterns and presence in areas beyond national jurisdiction, many highly migratory species are managed by Regional Fisheries Management Organizations (RFMOs).

@@

Many transboundary species are not considered highly migratory but are still shared by numerous neighboring EEZs (e.g. garfish *Belone belone*) (Fig. 4a). In addition, many fish stocks have meta-populations that are connected through larvae dispersal with ‘source’ populations potentially supporting ‘sink’ populations thousands of kilometers away [@Ramesh:2019va]. For example, while coral reef-related species were found to share fewer neighboring EEZs than other species (Fig. 4a), coral regions often share multiple species through larval connectivity [@Schill:2015fh] and adult movement [@Levin:2018in]. However, **it is important to acknowledge the uncertainty in the connectivity of marine populations at different life stages from larvae [@Kaplan:2016hs] to adults [@Archambault:2016ie].** In some areas like the Caribbean and the Western Indian Ocean, among other regions, transboundary marine protected areas have been identified as potential tools to support fisheries and conservation goals [@Levin:2018in; @Perez:2019dm]. The effective management of coral reef species is critical to many coastal communities, as they tend to highly dependent on subsistence fishing for food and nutrition security as well as livelihoods [@CisnerosMontemayor:2016gq; @Hanich:2018dh]. For instance, a number of countries with the largest catches of transboundary reef and pelagic-oceanic associated species (Fig. 4b) are also associated with some of the highest rates of fish consumption [@FAO:2018KK]. In the Philippines both pelagic and reef fishes contribute substantially to both local food and nutrition security as well as livelihoods [@Cabral:2018hc].

@@

The *Sea Around Us* has reconstructed global fisheries catches from 1950 to 2014, identifying commercial marine fish and invertebrates as well as fishing regions [@Zeller2016]. We used the *Sea Around Us* reconstructed species list to determine the number of transboundary marine species exploited by fisheries within each of the world's EEZs. The total number of species analyzed was 938, representing 67% of identified species by the *Sea Around Us* and accounting for 96.52% of the catch identified at the species’ level. To determine the current distribution of exploitable marine species, we used four data sources of species-distributions: (*i*) occurrence data, (*ii*) an Ensemble Environmental Niche Model (ENMs), (*iii*) a life-history-based distribution model, and (*iv*) fisheries catch data (Table 1). Each source represents a different method of estimating the distribution of a given species, thus providing a more robust result than an analysis focused on a single data source. Only commercial fished species with data from all four sources were included in our analysis.

@@

The SDM-SAU model follows a five-steps process based on species-specific life history data, rather than environmental variables [@Close:2006ux; @Pauly:2016uz]. For each commercial marine species, the model first uses the FAO major fishing areas and countries' EEZs to determine a broad distribution. It then uses life history information to delimit its range within the FAO fishing area (e.g., thermal preference, depth limit). The range is delimited even further by expert-review polygons and compared with that of AquaMaps [@Kaschner:2016tl], OBIS and GBIF occurrence data. The model then determines a species' habitat-preference based on the assumption that the relative abundance of a species is determined by the number of habitats in a grid cell and the species' distance to each habitat, as well as the importance of the habitat to the total size of the species distribution. Finally, the species equatorial submergence (e.g., the latitudinal region where a species is not seen in between poles) is estimated for each species. See [@Close:2006ux; @Pauly:2016uz] for model details.

>>>>

Regulatory boundaries and species distributions often do not align. This is especially the case for marine species crossing multiple Exclusive Economic Zones (EEZs). Such movements represent a challenge for fisheries management as policies tend to focus at the national level, yet international collaborations are needed to maximize long-term ecological, social and economic benefits of shared marine species. Here, we combined species distributions and the spatial delineation of EEZs at the global level to identify the number of commercially exploited marine species that are shared between neighboring nations. We found that 67% of the species analyzed are transboundary (n = 633). Between 2005 and 2014, fisheries targeting these species within global-EEZs caught on average 48 million tonnes per year, equivalent to an average of USD 77 billion in annual fishing revenue. For select countries, over 90% of their catch and economic benefits were attributable to a few shared resources. Our analysis suggests that catches from transboundary species are declining more than those from non-transboundary species. Our study has direct implications for managing fisheries targeting transboundary species, highlighting the need for strengthened effective and equitable international cooperation.

@@

Distributions of marine species around the world are not constrained by human-made boundaries; rather they are shaped by biotic and abiotic factors as well as species' evolutionary history [@Hutchinson:1957nt; @Nelson:2016kk]. A species can be widely distributed (cosmopolitan) or endemic [@Nelson:2016kk]. Fisheries management is predicated on the definition of “stocks”, delineated, in most cases by human-made spatial boundaries that often do not correspond to biologically-meaningful population units [@Song:2017iua; @FredstonHermann:2018kp]. The **development** of Exclusive Economic Zones (EEZs) under Part V of the United Nations Convention on the Law of the Seas (UNCLOS) in the early 80s [@UnitedNations:1986tl], for instance, extended political boundaries from 12 to 200 nautical miles to give coastal nations property rights over marine resources [@osthagen:2020hd]. However, EEZ boundaries cut across the distribution of many species, creating shared stocks between nations [@UnitedNations:1986tl]. Shared stocks can be classified into three non-exclusive categories; "transboundary" - stocks that cross the EEZs of two or more bordering coastal states; "straddling" - stocks that cross neighboring EEZs and the adjacent high seas; and "highly migratory" - stocks that cross non-neighboring EEZs and the high seas (mainly tunas) [@Munro:2004th]. Our study focuses on the "transboundary" nature of shared species exploited by fisheries operating within EEZs.

Theory and empirical evidence have shown that fisheries targeting resources that straddle political boundaries complicate fisheries management and potentially reduce the effectiveness of policies to achieve their stated objectives [@Munro:2002uf; @Englander:2019dq]. For instance, climate-driven changes in species distributions have led to conflicts between nations, driven, at least partly, by changes in the proportion of captures [@Munro:2002uf], quota allocation, and fishery newcomers [@Spijkers:2017ij]. Moreover, climate change is likely to exacerbate such conflicts and presents new challenges for political relations between neighboring countries [@Pinsky:2018cb] and fisheries management [@Miller:2013iv]. Therefore, having an accurate understanding of the distribution and scale of transboundary and straddling fish stocks as well as associated fisheries is important to inform their sustainable management, particularly under climate change.

Forty years after the formal adoption of UNCLOS [@UnitedNations:1986tl] and the subsequent 1995 United Nations Fish Stock Agreement for the cooperation on the management and conservation of straddling and highly migratory resources [@UnitedNations:1995fs], accurate estimates of the number of exploited marine species shared by neighboring nations are still unavailable. An informed guess, based on limited biogeographical data, suggested that there are approximately 500 to 1500 exploited transboundary stocks [@Caddy:1997ue]. A recent literature review included 344 shared taxa and their historical contributions to fisheries [@Teh:2015gd]. However, these studies did not consider species' biogeography in the quantification of transboundary stocks. Here, we aim to estimate the number of exploited marine species shared by neighboring countries and determine their contribution to global and regional catches as well as fisheries revenue. Moreover, we categorize species according to their catch trends while identifying differences among species based on habitat preference. We hypothesize that the methodological constraints of previous studies resulted in an underestimation of the number of transboundary species and their contribution to global catch and revenue.

We overlaid the known distribution of 938 commercially valuable marine species responsible for an average of 96.5% of global EEZ catches between 2005 and 2014, and 280 EEZs of 198 coastal countries (See methods). We define a ‘stock’ unit as a species in an EEZ, instead of a genetically or morphological distinct unit [@Nelson:2016kk], due to the lack of such biological information being available for almost all the species included in this study [@Teh:2015gd]. While we acknowledge that species could have multiple stocks within an EEZ, many fisheries within a country or EEZ are managed at the species instead of stock level (or even as groups containing multiple species). For example, shrimp (e.g., *Litopenaeus stylirostris* or *Farfantepenaeus californiensis*) along the Pacific coast of Mexico [@DiarioOficialdelaFederacionDOF:2018wt] and hammerhead shark (*Sphyrna zygaena*) in Peru [@MAP:2017uf] are managed at the species level, yet include multiple populations exploited by each country’s fisheries. Moreover, recent research shows connectivity across fish stocks through larval dispersal [@Ramesh:2019va] and adult migration [@Levin:2018in; @Popova:2019fs; @Dunn:2019ca], although considerable level of uncertainty exists at different life stages [@Kaplan:2016hs; @Archambault:2016ie]. For this analysis we only considered shared species between neighboring EEZs, rather than the species’ extended distribution (e.g., we did not include the high seas). We rely on multiple data sources including occurrence, distribution models and catch data, and only consider a species to be present in a grid cell if all data sources showed positive occurrence (see Methods).

@@

We identified 633 exploited transboundary species worldwide (67.5% of the 938 species analyzed), almost double previous estimates [@Teh:2015gd]. Between 2005 and 2014, national fleets targeting these transboundary species within EEZs caught an annual average of 48.5 million tonnes, representing 82.3% of EEZ-derived catches reconstructed by the *Sea Around Us* at the species level (Fig. 1a). These catches generated a yearly average of USD 77,591 million in fishing revenue (78.5% of global fishing revenue) over the same time period. Our findings are considerably higher than the previous estimates of 34.2 million tonnes in catches and USD 40,187 million (in 2019 value) in fishing revenue from shared stocks [@Teh:2015gd]. When we re-estimated transboundary species’ catches using data consistent with those used in previous studies (FAO global reported data in year 2006), the 633 transboundary species identified here accounted for 40.4 million tonnes of annual catches (i.e., 18% higher than previously estimated). Our results suggests that the contribution of transboundary species to global catch and fishing revenue might previously have been  underestimated due to an incomplete understanding of the transboundary nature of marine fished species. The 305 non-transboundary species (termed as ‘discrete’ species here, see methods - Determining transboundary species trait) accounted for a much smaller proportion of total catch and revenue; 2.8 million tonnes and USD 4,282 million annually respectively, on average, between 2005 and 2014. These results underscore the importance of transboundary species at the global level.

@@

In many cases, according to our categorization criteria, a transboundary species **can be distributed** in multiple EEZs but only counted as transboundary in a subset of EEZs (see Methods). For example, the distribution of Peruvian anchoveta (*Engraulis ringens*) spans the EEZs of Peru, Chile, and Ecuador. However, our study only considered the stocks in Peru and Chile as transboundary [@Cashion:2018cg]. Anchoveta in Ecuadorian waters only include a small proportion of the shared distribution range (spatial threshold between Ecuador and Peru < 10%; see methods - Criteria 3) and thus do not meet our criteria for consideration as a transboundary stock. A situation similar to the example of Peruvian anchoveta in Ecuadorian waters is common amongst the transboundary species identified in this study. Overall, 590 of the 633 transboundary species have stocks in EEZs that do not meet our criteria for consideration as transboundary stock. The annual average contribution from all excluded stocks of transboundary species to fisheries (e.g., Ecuador’s anchoveta catch) between 2005 and 2014 was 10.8 million tonnes, representing USD 19,243 million in fishing revenue over the same time period.

@@

Peru and Russia, having the two largest fisheries by total catch in the world [@FAO:2018KK], were each responsible for over 5.8 million tonnes of transbounuadry species catch annually, on average, between 2005 and 2014  (Supplementary Fig. S1). Peru's catches consisted mainly of Peruvian anchoveta (*Engraulis ringens*) and accounted for 79% of the national transboundary species production. Peru and Chile recently signed an agreement to work towards standardized stock assessments through coordinated management of the southern anchoveta stock [@UNDP:2016wp]. A similar management agreement was signed by Russia, Japan and the USA over shared Alaskan pollock (*Theragra chalcogramma*) in the Bering Sea in 1988 [@NOAA:gX6CtUTk]. Transboundary species also make large contributions to fisheries in Eastern Asia (one-way ANOVA; *Df*(16,158) = 2.265, *p* = 0.005, $\alpha$ = 0.05). China, the world's top fish producer [@FAO:2018KK], obtains one third (5.1 million tonnes) of its total catches from transboundary species, followed by Japan (1.8 million tonnes) and South Korea (1.06 million tonnes). Differences in the regional importance of transboundary fisheries can also be illustrated in terms of catch-revenue over the area ($km^2$) of the EEZ (Fig. 2). As an example, the aggregated EEZ area for all Northern European countries where transboundary species are present is 3.3 million $km^2$, the $6^{th}$ smallest of the 17 groups analyzed (Supplementary Table S1). However, it had the second highest average revenue (USD 26.1 thousand per $km^2$) and the highest average catch (19.9 tonnes per $km^2$) of transboundary species per EEZ area between 2005 and 2014. At the country level, countries from Western Europe accrued significantly more revenue from transboundary fisheries per $km^2$ than any other country (one-way ANOVA, *Df*(16,165) = 3.267, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* $\leq$ 0.05; Fig. 2).

@@

![Figure 2. Weighted benefits of transboundary species by $km^{2}$ and UN sub-region. a) Revenue in thousand 2019 USD. b) Catch in tonnes. Points = sub-region mean $\pm$ s.d. by country. Figure created using R version 3.5.2. Abbreviations as in Figure 1.](../Figures/Fig2.png)

We determined the catch trend of each species within each EEZ, classifying them as increasing (Category A), constant (Category B) or decreasing (Category C) (Fig. 3,  Supplementary Fig. S2). While previous studies have demonstrated that catch trends can be used to infer whether a stock is healthy, re-building, over-exploited or collapsed [@Kleisner:2011wn], several factors can influence stock status. Our intention here is to examine where the catch trends of transboundary species differ from non-transboundary species between 2005 and 2014 relative to historic catch since 1951 (see methods). We found significant differences for all catch trend categories for transboundary species (one-way ANOVA, *Df*(2,459) = 47.94, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* $\leq$ 0.001), and no significant differences in catch trend categories for discrete species (one-way ANOVA, *Df*(2,106) = 1.885, *p* = 0.157; $\alpha$ = 0.05). We also found significant differences in catch trends when directly comparing transboundary to discrete species categories (MANOVA, *Df*(2,459) = 19.001, *p* < 0.001). Overall, transboundary species only targeted by one country are generally less likely to have a decreasing catch trend compared to instances where the shared species is fished by neighboring countries (Supplementary Table S2).

@@

![Figure 3. Number of transboundary species by catch trend and EEZ. Category A, Increasing; Category B, Constant; Category C, Decreasing. “No Category” reflects species with less than 10 years of catch data and/ or less than 5 consecutive years of catch data. Only showing first 100 species. Figure created using R version 3.5.2.](../Figures/Fig3.png)

Empirical analysis suggests that in most cases, management of transboundary species will yield better outcomes in terms of fish catches when nations cooperate [@Munro:2002uf]. Yet, cooperation can be a complex process [@Jensen:2015cf], and in specific cases joint management might not be the best strategy [@Munro:2015df]. Examples of successful joint management include agreements between Norway and Russia over Atlantic cod (*Gadus morhua*) [@Eide:2013dha] and Namibia and South Africa over hake (*Merluccius spp*) [@Sumaila:2003vw]. Lack of collaboration over shared stocks may threaten stock sustainability, reduce the future profitability potential of the fishery, and result in conflict between neighboring nations [@Clark:1980js; @Spijkers:2017ij].

Transboundary fisheries are important to a number of countries with notorious fisheries-related conflicts, including Canada, the USA, the European Union (EU) and Russia [@Spijkers:2019cz]. For example, since 2007, the EU, Norway, Iceland, and the Faroe Islands (Denmark) have been at odds over the size and relative allocation of the total allowable catch for Atlantic mackerel (*Scomber scombrus*) due to the species' climate-driven shift in distribution [@Spijkers:2017ij]. Atlantic Mackerel is a transboundary species that straddles into the high seas. Among the countries involved in the 2007 fisheries dispute, Atlantic mackerel contributed an annual average (between 2005 and 2014) of 598.4 thousand tonnes (8.19%) in total catch and USD 850.2 million (7.17%) in total fishing revenue. Climate change is expected to continue changing the distribution and shared proportion of fish stocks among countries, resulting in the emergence of new transboundary species [@Pinsky:2018cb], and disappearance of some species from EEZs [@Oremus:2020id]. Exploring the detailed effects of climate change on the distribution of shared species is key to the development of local adaptation methods that can anticipate negative impacts to sustainability. For example, understanding how climate change will modify the proportion of transboundary species shared by neighboring EEZs, the time-frame over which such changes will happen [@PalaciosAbrantes:2020vt], and the economic consequence of such effects can inform the development of more anticipatory and climate-resilient international treaties improving fisheries management [@Sumaila:2020tp].

Most marine fish species occur in tropical and subtropical waters around the world [@Nelson:2016kk; @Reygondeau:2019uh], from highly migratory species associated with pelagic-oceanic ecosystems like tunas (*Thunnus sp.*), to less mobile reef-associated species like greater amberjack (*Seriola dumerili*), and species found in demersal ecosystems like gilthead seabream (*Sparus aurata*). Species associated with pelagic-oceanic ecosystems are the only group whose EEZ range (i.e., the number of EEZs where the species occur as transboundary) is significantly different from other groups (one-way ANOVA; *DF*(5,597) = 53.82, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05), with a median of 40 EEZs per species. The median for species of all other ecosystem preferences is close to, or less than, 20, as many of these species have a narrower distribution or are less mobile (Fig. 4a). This result is likely an effect of the broad distribution that many of the species with preference for pelagic-oceanic ecosystems have. Many pelagic-oceanic species are highly migratory and straddle across EEZs while crossing the high seas. Thus, the number of non-neighboring EEZs sharing a highly migratory species can be over 100, as in the case of bigeye tuna (*Thunnus obesus*) (Fig. 4a). Due to their vast migration patterns and presence in areas beyond national jurisdiction, many highly migratory species are managed by Regional Fisheries Management Organizations (RFMOs).

@@

Many transboundary species are not considered highly migratory, but are still shared by numerous neighboring EEZs (e.g., garfish *Belone belone*) (Fig. 4a). In addition, many fish stocks have meta-populations that are connected through larval dispersal with ‘source’ populations potentially supporting ‘sink’ populations thousands of kilometers away [@Ramesh:2019va]. For example, while coral reef-associated species were found to share fewer neighboring EEZs than other species (Fig. 4a), coral regions often share multiple species through larval connectivity [@Schill:2015fh] and adult movement [@Levin:2018in]. However, it is important to acknowledge the uncertainty in the connectivity of marine populations at different life stages from larvae [@Kaplan:2016hs] to adults [@Archambault:2016ie]. In some areas like the Caribbean and the Western Indian Ocean, among other regions, transboundary marine protected areas have been identified as potential tools to support fisheries and conservation goals [@Levin:2018in; @Perez:2019dm]. The effective management of coral reef species is critical to many coastal communities, as they tend to be highly dependent on subsistence fishing for food and nutrition security as well as livelihoods [@CisnerosMontemayor:2016gq; @Hanich:2018dh]. For instance, a number of countries with the largest catches of transboundary reef and pelagic-oceanic associated species (Fig. 4b) are also associated with some of the highest rates of fish consumption [@FAO:2018KK]. In the Philippines, both pelagic and reef fishes contribute substantially to both local food and nutrition security, as well as livelihoods [@Cabral:2018hc].

@@

The *Sea Around Us* has reconstructed global fisheries catches from 1950 to 2014, identifying commercial marine fish and invertebrates as well as fishing regions [@Zeller2016]. We used the *Sea Around Us* reconstructed species list to determine the number of transboundary marine species exploited by fisheries within each of the world's EEZs. The total number of species analyzed was 938, representing 67% of identified species by the *Sea Around Us* and accounting for 96.52% of the catch identified at the species’ level. To determine the current distribution of exploitable marine species, we used four data sources of species-distributions: (*i*) occurrence data, (*ii*) an Ensemble Environmental Niche Model (ENMs), (*iii*) a life-history-based distribution model, and (*iv*) fisheries catch data (Table 1). Each source represents a different method of estimating the distribution of a given species, thus providing a more robust result than an analysis focused on a single data source. Only commercially fished species with data from all four sources were included in our analysis.

@@

The SDM-SAU model follows a five-step process based on species-specific life history data, rather than environmental variables [@Close:2006ux; @Pauly:2016uz]. For each commercial marine species, the model first uses the FAO major fishing areas and countries' EEZs to determine a broad distribution. It then uses life history information to delimit its range within the FAO fishing area (e.g., thermal preference, depth limit). The range is delimited even further by expert-review polygons and compared with distributions from AquaMaps [@Kaschner:2016tl], as well as OBIS and GBIF occurrence data. The model then determines a species' habitat-preference based on the assumption that the relative abundance of a species is determined by the number of habitats in a grid cell and the species' distance to each habitat, as well as the importance of the habitat to the total size of the species distribution. Finally, the species equatorial submergence (e.g., the latitudinal region where a species is not seen in between poles) is estimated for each species. See [@Close:2006ux; @Pauly:2016uz] for model details.
